Field-informed Reinforcement Learning for Learnining Large Scale Collective Tasks

### 2023-04-26 paper.tex
chore: more on many agent rl

\acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}

\begin{abstract}
Coordinatinating a group of intelligent agents in multi-agent systems 
 is a research problem that has been addressed for a long time, 
 due to the challenges posed by distribution and of the definition of distributed intelligence. 
%
The problem is even more evident in collective adaptive systems, 
 where the scale of the systems considered makes the definition of collective behaviors even more challenging.

@@

Reinforcement learning (RL) has gained a lot of interest recently, 
 thanks to its successful application in various scenarios, 
 ranging from video games (such as Alpha Go and Atari) 
 to chatbots (like ChatGPT). 
% 
In RL, an agent (i.e., a smart entity capable of making decisions) 
 performs actions in an environment (i.e., everything outside the agent) according to a policy, 
 with the goal of maximizing long-term rewards.

One interesting application of RL is when there are multiple learning agents involved. 
 Such scenarios are referred to as multi-agent reinforcement learning (MARL). 
 While standard RL approaches use Markov decision processes to model the world, 
 in MARL for distributed systems, the notion of 
 \acp{DecPOMDP} is used.

A \ac{DecPOMDP} represents a sequential game, 
 in which each agent receives a local observation about the world they 
 inhabit and receives a joint immediate reward in response to the actions they take. 
% 
\ga{Improve, if necessary, the following paragraph.}
Formally, a DecPOMDP is defined as a tuple $<N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma>$, 
 where $N$ is the total number of agents that live in the environment, 
 $S$ is the state set of the environment, 
 $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
 $P$ is the probability transition function, 
 $R$ is the immediate joint reward function, and 
 $\{\Omega_i\}$ is the set observation function for the agent $i$,
 $O$ is the observation space, and finally $\gamma$ is the discount factor.
%
On top of this model, each agent try to learn a policy $\pi_i$ in order 
 to maximise the long term join reward signal.

In base of some of this concept, 
 we have several different MARL, like cooperative MARL (i.e., the reward function incetivize collective outcomes)
 heterogenous MARL (i.e., the agents have different action space and policy) and
 homogenous MARL (i.e., the agents have the same action space and policy).
%  
In particular, in this work we consider homogenous many agent RL, 
 where, N >> 2 and action space of each agent and the policy is the same (i.e., the agents are interchangeable)
\section{Field-informed Reinforcement Learning}
\label{sec:approach}
\ga{Page budget: 2/3 pages \\}
\ga{Plan: we should discuss the approach in detail, including the system model, the learning dynamics.
Particularly, I would like to highlight how the each node in the graph is a local controller,
but it could be see as a global evolution, therefore we can use global information to inform the local controller.}
\subsection{System model}
\ga{Plan: 
  discussion about typical AC evalaution model, specifying how GNN (with 1-hop information diffusion)
  can be easily integrate with this distributed model
}
\subsection{Learning dynamics}
\ga{
  Plan: discuss about the choosing approach, that is Centralised Training and Distributed Execution.
  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consist in a set of graphs,
  where each graph is a snapshot of the system. 
  The target is the Q-value of the action taken in the current state.
  Explain how this can be then use for distributed controller.
}
\section{Evaluation}
\label{sec:eval}
\ga{Page budget: 2 pages \\}
\ga{Plan: we should discuss the robot aggregation scenario, specifiying the state space, the action space, the reward function.
I dunno if it makes sense to discuss the variants (i.e., fixed area, two ares and moving area). We will see.
}
\subsection{Scenario}

>>>>

\acrodef{DecPOMDP}[DecPOMDP]{decentralized partially observable Markov decision processes}
\acrodef{RL}[RL]{reinforcement learning}
\acrodef{MARL}[MARL]{multi-agent reinforcement learning}
\begin{abstract}
Coordinatinating a group of intelligent agents in multi-agent systems 
 is a research problem that has been addressed for a long time, 
 due to the challenges posed by distribution and of the definition of distributed intelligence. 
%
The problem is even more evident in collective adaptive systems, 
 where the scale of the systems considered makes the definition of collective behaviors even more challenging.

@@

\Ac{RL} has gained a lot of interest recently, 
 thanks to its successful application in various scenarios, 
 ranging from video games (such as Alpha Go~\cite{Silver2016Go} and Atari~\cite{Atari2016DQN}) 
 to chatbots (like ChatGPT~\cite{ChatGPT2023}). 
% 
In \ac{RL}, an agent (i.e., a smart entity capable of making decisions) 
 performs actions in an environment (i.e., everything outside the agent) according to a policy, 
 with the goal of maximizing long-term rewards.

One interesting application of \ac{RL} is when there are multiple learning agents involved. 
 Such scenarios are referred to as multi-agent reinforcement learning (MARL)~\cite{zhang2019marl}. 
 While standard \ac{RL} approaches use Markov decision processes to model the world, 
 in MARL for distributed systems, the notion of 
 \acp{DecPOMDP}~\cite{Decpomdp2000} is used.

A \ac{DecPOMDP} represents a sequential game, 
 in which each agent receives a local observation about the world they 
 inhabit and receives a joint immediate reward in response to the actions they take. 
% 
\ga{Improve, if necessary, the following paragraph.}
Formally, a \ac{DecPOMDP} is defined as a tuple $\langle N, S, \{A_i\}, P, R, \{\Omega_i\}, O, \gamma \rangle$, 
 where $N$ is the total number of agents that live in the environment, 
 $S$ is the state set of the environment, 
 $\{A_i\}$ is the set of actions for agent $i$ and $A$ is the joint action space, 
 $P$ is the probability transition function, 
 $R$ is the immediate joint reward function, and 
 $\{\Omega_i\}$ is the set observation function for the agent $i$,
 $O$ is the observation space, and finally $\gamma$ is the discount factor.
%
On top of this model, each agent try to learn a policy $\pi_i$ in order 
 to maximise the long term joint reward signal.
In base of some of these concepts, 
 there are several approaches/classifications in literature, 
 like cooperative (i.e., the reward function incetivize collective outcomes)
 heterogenous (i.e., the agents have different action space and policy) and
 homogenous (i.e., the agents have the same action space and policy) settings.
%  
In particular, in this work we consider homogenous many agent RL~\cite{yang2021many}, 
 where $N \gg 2$ and each agent is interchangeable and indistinguishable.
%
This research area is relevant in the context of in large-scale systems 
 when the incentive of the individual is reduced compared to the collective behavior of the system, 
 which is found in collective adaptive systems such as swarm robotics.

In fact, the first works in this direction were discussed precisely in the last field, 
 exploring new models (e.g., swarMDP~\cite{DBLP:conf/atal/SosicKZK17}) and learning algorithms capable of extrapolating 
 a policy equal for the entire system. 
Modern approaches, however, 
 have started to consider the use of deep learning approaches to try 
 to synthesize robust controllers capable of generalizing to new tasks. 
In this context, mean-field reinforcement learning~\cite{pmlr-v80-yang18d} is certainly noteworthy. 
 Mean-field RL is a technique for learning approximate inference in large-scale systems, 
 where the interactions between the agents can be decoupled into a 
 single conditional probability distribution function for each agent, 
 thus simplifying the learning process and reducing the computational cost.
%
Some known approaches using mean-field reinforcement learning include Q-mean, 
 which is an extension of Q-learning to mean-field settings, 
 and actor-critic mean-field, which combines actor-critic algorithms with mean-field approximations. 
%
These approaches have shown promising results in various domains, such as multi-agent coordination 
 and decentralized control, and are actively being researched and developed for further applications

\section{Field-informed Reinforcement Learning}
\label{sec:approach}
\ga{Page budget: 2/3 pages \\}
\ga{Plan: we should discuss the approach in detail, including the system model, the learning dynamics.
Particularly, I would like to highlight how the each node in the graph is a local controller,
but it could be see as a global evolution, therefore we can use global information to inform the local controller.}
\subsection{System model}
\ga{Plan: 
  discussion about typical AC evalaution model, specifying how GNN (with 1-hop information diffusion)
  can be easily integrate with this distributed model
}
\subsection{Learning dynamics}
\ga{
  Plan: discuss about the choosing approach, that is Centralised Training and Distributed Execution.
  Particularly, we use DQN in which the neural network is a GNN. So, the dataset consist in a set of graphs,
  where each graph is a snapshot of the system. 
  The target is the Q-value of the action taken in the current state.
  Explain how this can be then use for distributed controller.
}
\section{Evaluation}
\label{sec:eval}
\ga{Page budget: 2 pages \\}
\ga{Plan: we should discuss the robot aggregation scenario, specifiying the state space, the action space, the reward function.
I dunno if it makes sense to discuss the variants (i.e., fixed area, two ares and moving area). We will see.
}
\subsection{Scenario}
